---
title: "ggplot_notes_Ben"
author: "Ben Kubacki"
date: "8-30-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### ggplot practice

textbook chapter 3: https://djvill.github.io/r4ds/data-visualisation.html

#### Pre-req Load Tidyverse
```{r 3.1 Prerequisite - load tidyverse package every time}
library(ggplot2)
library(tidyverse)

```

## 3.2 First Steps

What is the relationship between engine size and fuel efficiency?

My hypothesis: Intuitively, engine size could mean a more powerful engine, thus better at using fuel. Thus, a larger engine could use fuel more efficiently. However, I know from experience/general knowledge that larger and heavier duty vehicles use fuel less efficiently, and I imagine such vehicles require larger engines, so I think it's safer to guess that larger engines have lower fuel efficiency. On the graph, this would look like a line that slopes downward as engine size (x = displ) increases, meaning larger engines travel less miles per gallon (y = hwy). As miles per gallon increases (y-axis), engine size decreases.

`displ` = engine size in liters
`hwy` = efficiency in miles per gallon

```{r mpg summary & scatterplot}
ggplot2::mpg

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

```
The graph supports our hypothesis (honestly, I'd seen the graph already). Thus, it supports that fuel efficiency decreases as engine size increases; the two variables are inversely proportional. 

### 3.2.3 Graphing template

`ggplot(data = <DATA>) +   
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))`
  
#### 3.2.4 Exercises

1. Running `ggplot(data = mpg)` delivers a blank graph, in fact it's just a gray box.

2. There are 234 rows and 11 columns in the dataset. You can use nrow/ncol or `ggplot2::mpg` to get the tibble numbers.

3. `drv` is the type of drive train. `f` = front-wheel, `r` = rear-wheel, and `4` = 4wd.

4. Scatterplot of x = cyl and y = hwy.

5. Scatterplot of `class` vs `drv` is not very useful as as scatter because it simply describes what type of drive each class of vehicle has. So we find that SUVs may have 4- or rear-wheel drive but not front-wheel drive. A scatterplot is meant to display numeric data relationships to numeric or nominal variables, rather than compare two nominal variables.

```{r 3.2.4 Exercises}

# ggplot(data = mpg)

nrow(mpg)
ncol(mpg)
ggplot2::mpg

# ?mpg

# scatterplot of hwy and cyl

# `cyl` = number of cylinders 

ggplot(data = mpg) + geom_point(mapping = aes(x = cyl, y = hwy))

ggplot(data = mpg) + geom_point(mapping = aes(x = drv, y = class))


```

### 3.3 Aesthetics

```{r}

ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, color = class))

# This suggests 2-seaters are unusually high mpg for engine size.

```
So we can use 

- color
- size
- shape (6 variables) 
- alpha (shading) 

to display variables by aesthetic **levels**. 

### 3.3.1 Exercises

1. Setting a color manually for your output needs to go outside the aes() argument. The code was wrong because the `color` argument was inside the aes() argument, which led to R interpreting it as a variable with a single level called blue. This is the correction: 

`ggplot(data = mpg)+geom_point(mapping=aes(x=displ, y=hwy), color = "blue")`

2. Categorical variables: 